---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quarantinesim

<!-- badges: start -->
<!-- badges: end -->

`quarantinesim` provides a ready-to-use dataset of quarantine outbreak
scenarios that can be explored through both R and a built-in Shiny app.

The dataset has already been merged and cleaned, and each row represents
a simulated epidemiological scenario defined by:

- **R0** : basic reproduction number  
- **VE** : vaccine effectiveness  
- **coverage** : vaccination coverage proportion  

It also includes modelled outputs such as:

- `traveller_ob_prob`: outbreak probability from travellers  
- `worker_ob_prob`: outbreak probability from quarantine workers  
- `chance50`: time or iteration when outbreak probability reaches 50%  
- `chance95`: time or iteration when outbreak probability reaches 95%  

The package is designed for **exploratory data analysis and visualisation**, and it includes a **Shiny app** that allows users to interactively explore how outbreak risk changes with **coverage, R0, and VE**, and to switch between different outcomes (probabilities and time-to-outbreak).

## Installation

You can install the development version of `quarantinesim` from GitHub with:

```r
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-Melisa-ops")
```


Or, if you are working locally inside the package project:
```r
install.packages("devtools")  # if not installed yet

devtools::install()
```

Then load the package:
```r
library(quarantinesim)
```

The app lets you:

- select a vaccine coverage,
- choose which outcome to display (traveller, worker, chance50, chance95),
- and switch chart type (line, point, bar), while also viewing the filtered data in a separate tab.


## Example

```{r}
library(quarantinesim)

# load dataset shipped with the package
data("data_quarantine")

# see sample of data
head(data_quarantine)

# outbreak probability by R0 for VE =0.5
data_quarantine |>
  dplyr::filter(VE == 0.5) |>
  ggplot2::ggplot(ggplot2::aes(x = R0)) +
  ggplot2::geom_line(ggplot2::aes(y = traveller_ob_prob, colour = factor(coverage))) +
  ggplot2::geom_line(
    ggplot2::aes(y = worker_ob_prob, colour = factor(coverage)),
    linetype = "dashed"
  ) +
  ggplot2::labs(
    y = "Outbreak probability",
    colour = "Vaccine coverage",
    title = "Outbreak probability by R0 (VE = 0.5)",
    subtitle = "solid = travellers, dashed = workers"
  ) +
  ggplot2::theme_minimal()


```

## Launch the Shiny app
To explore the data interactively, run:

```r
run_quarantinesim()
```

(Make sure you have inst/app/app.R and an exported run_quarantinesim() function in R/.)

You can also view the app source files on GitHub:
https://github.com/ETC5523-2025/assignment-4-packages-and-shiny-apps-Melisa-ops/tree/main/quarantinesim/inst/app

## Overview of components

- data_quarantine : final dataset containing merged outbreak scenarios
- run_quarantinesim() : launches the interactive Shiny app
- Shiny app in inst/app/ : UI to explore outbreak probabilities across R0, VE, and coverage
- Package documentation : created with roxygen2 for data and functions
